---
title: "share_analysis"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fundamental and Technical Analysis of Shares 

[Original Source](http://www.r-exercises.com/2016/08/31/fundamental-and-technical-analysis-of-shares-exercises/)

Explore possibilities for fundamental and technical analysis of stocks offered by the **quantmod** package.

```{r}
library(quantmod)
```


## Exercise 1

Load FB (Facebook) market data from Yahoo and assign it to an xts object fb.p.

```{r}
fb.p <- getSymbols("FB", env=NULL)
```

## Exercise 2

Display monthly closing prices of Facebook in 2015.

```{r}
Cl(to.monthly(fb.p["2015::2015-12-31"]))
```

## Exercise 3

Plot weekly returns of FB in 2016.

```{r}
plot(weeklyReturn(fb.p, subset="2016::"), main="Weekly return of Facebook")
```

## Exercise 4

Plot a candlestick chart of FB in 2016.

```{r}
candleChart(fb.p, subset="2016::2016-12-31", name="Facebook", theme="white")
```

## Exercise 5

Plot a line chart of FB in 2016., and add boilinger bands and a Relative Strength index to the chart.

```{r}
chartSeries(fb.p, subset="2016::2016-12-31", type="line", name="Facebook", theme="white")
addBBands()
addRSI()
```

## Exercise 6

Get yesterday's EUR/USD rate.

```{r}
getFX("EUR/USD", from=Sys.Date()-1, env=NULL)
```

## Exercise 7

Get financial data for FB and display it.

```{r}
fb.f <- getFin("FB", env=NULL)
viewFin(fb.f)
```

## Exercise 8

Calculate the current ratio for FB for years 2013, 2014 and 2015. (Tip: You can calculate the current ratio when you divide current assets with current liabilities from the balance sheet.)

```{r}
fb.bs <- viewFin(fb.f, "BS", "A")
fb.bs["Total Current Assets", c("2013-12-31", "2014-12-31", "2015-12-31")] / fb.bs["Total Current Liabilities", c("2013-12-31", "2014-12-31", "2015-12-31")]
```

## Exercise 9
 
Based on the last closing price and income statement for 12 months ending on December 31th 2015, Calculate the PE ratio for FB. (Tip: PE stands for Price/Earnings ratio. You calculate it as stock price divided by diluted normalized EPS read from income statement.)

```{r}
price <- Cl(fb.p[NROW(fb.p)])
fb.is <- viewFin(fb.f, "IS", "a")
EPS <- fb.is["Diluted Normalized EPS", "2015-12-31"]

price / EPS
```

## Exercise 10

write a function getROA(symbol, year) which will calculate return on asset for given stock symbol and year. What is the ROI for FB in 2014. (Tip: ROA stands for Return on asset. You calculate it as net income divided by total asset.)

```{r}
getROA <- function(symbol, year) {
  symbol.f <- getFin(symbol, env=NULL)
  symbol.ni <- viewFin(symbol.f, "IS", "A")["Net Income", paste(year, sep="", "-12-31")]
  symbol.ta <- viewFin(symbol.f, "BS", "A")["Total Assets", paste(year, sep="", "-12-31")]
  symbol.ni / symbol.ta * 100
}

getROA("FB", 2014)
```

